Below-ground sink limitation alters growth and carbon balance of Eucalyptus seedlings

### 2015-10-24 manuscript/manuscript_pve.Rmd
manuscript edits from MT

\
We then utilized this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf A where sufficient to accurately predict overall seedling biomass production after 120 days. Each model run utilized changes in A and leaf mass fraction (LMF), with values of stem and root respiration rates, to generate total seedling mass and leaf area after 120 days.  Cumulative net leaf C gain for each treatment was equal to the sum of each value of C~day,i~ over 120 days and final seedling C was assumed to equal half of the final mass for both modeled and observed seedlings. First, a default model was optimized to produce a final LMF that correctly predicted both the final leaf mass and total biomass of the harvested free seedling controls (M~0~). This optimized LMF was then applied to model runs with treatment specific C~day~ to determine if changes in leaf A alone could accurately predict biomass (M~1~). Next, model sensitivity to different C allocation scenarios, including non-optimized treatment specific LMF and up regulation of non-leaf tissue respiration by 50 % of default values (M~2,3~, respectively), was used to improve predictions of initial model simulations from measured harvest biomass. For all cases, seedling biomass production was compared between model output and harvested seedlings with treatment specific mean daily C assimilation by first scaling values to the free seedling control.

@@

## Modelling seedling biomass
The default model M~0~, was optimized so that the free seedling would fit as well as possible and then used to predicted biomass for seedlings in containers. M~0~ was able to converge on an optimum LMF (21.6 %) which predicted the mean harvest total biomass of free seedlings within 1.2 %. Using this optimized LMF, the total biomass of modeled seedlings with soil volume restriction (M~1~) was on average 23&pm;2.4 g C more than measured seedlings when comparing against predicted total net leaf C gain (Figure `r massmodelfig`a). Thus, seedling C mass was overestimated by an average of 50&pm;8.7 % in modeled seedlings across the soil volume treatments (Figure `r massmodelfig`b). As a result, the observed reductions in leaf A with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the  container size treatments.

\
Next, we performed a series of model simulations to test possible C allocation scenarios to account for the over predictions of seedling C mass. Testing the sensitivity of the model to treatment-specific LMF from the final harvest (M~2~) improved model predictions of seedling C mass but still overestimated seedling total C by 32&pm;11.1 % (Figure `r massmodelfig2`a). Using harvest values of LMF, however, does not capture the increase in leaf turnover of seedlings in small containers (Figure `r allomfig`c). Thus, the 7-61 % reduction in harvest avlues of LMF from large to small soil volumes compared to the optimized LMF (M~0~) represents an underestimation of realized leaf C allocation. Increases of 50 % in non-leaf tissue respiration (M~3~) improved biomass estimates slightly but overestimated mass C by an average of 46&pm;9.3 % in seedlings with soil volume restriction (Figure `r massmodelfig2`b). With M~3~, non-tissue respiration rates would need to be increased by ca. 250% in order for mass balance to be achieved.

@@

\
Decreases in A~sat~ occurred at the same time as reductions in height and diameter of seedlings in containers. This initially suggests a strong link between growth and an apparent down regulation of A. However, there are several possible mechanisms that can explain reduced A in small pots including nutrient content, water or reduced sink strength [@poorter2012pot].  It was therefore necessary to examine each of these factors to determine if the induced belowground sink limitation actually triggered photosynthetic down regulation.

@@

## Conclusions
The debate over how rates of photosynthesis affect plant growth or to what degree these rates are instead controlled by growth has existed for decades [@sweet1966role]. Here, although biomass partitioning was conserved our model results infer that the allocation of photosynthate between non-biomass pools was altered by sink inhibition. This is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants has seldom been studied. Our findings add more evidence that A and growth are not always entirely synced, an important distinction often missed in studies that manipulate source/sink activity. K&#246;rner [-@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control A through the demand for C. Although this may be true, we argue that attempts to quantify or at least predict the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

>>>>

\
We then utilized this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf A where sufficient to accurately predict overall seedling biomass production after 120 days. Each model run utilized changes in A and leaf mass fraction (LMF), with values of stem and root respiration rates, to generate total seedling mass and leaf area after 120 days.  Cumulative net leaf C gain for each treatment was equal to the sum of each value of C~day,i~ over 120 days and final seedling C was assumed to equal half of the final mass for both modeled and observed seedlings.

First, a default model was fitted with mean photosynthetic parameters from the free seedlings and then optimized to produce a LMF that correctly predicted both the leaf mass and total biomass of the harvested free seedling (M~free~). This optimized LMF was then used to constrain model runs with soil volume treatment specific C~day~, while keeping tissue respiration rates and turnover parameters constant, to determine if changes in leaf A alone could predict biomass of seedlings in containers (M~pots~). Next, model sensitivity to different C allocation scenarios, including non-optimized treatment specific LMF and up regulation of non-leaf tissue respiration by 50 % of default values (M~s1,s2~, respectively), was used to improve predictions of model simulations compared to measured harvest biomass. For all cases, seedling biomass production was compared between model output and harvested seedlings with treatment specific mean daily C assimilation by first scaling values to the free seedling control.

@@

## Modelling seedling biomass
The default model M~free~, successfully optimized a LMF (21.6 %) which then allowed model to predict mean harvest total biomass of free seedlings within 1.2 %. Using this optimized LMF, the total biomass of modeled seedlings for each soil volume treatment (M~pots~) were on average 23&pm;2.4 g C more than measured seedling biomass when compared against predicted total net leaf C gain (Figure `r massmodelfig`a). Thus, seedling C mass was overestimated by an average of 50&pm;8.7 % in modeled seedlings across the soil volume treatments (Figure `r massmodelfig`b). As a result, the observed reductions in leaf A with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the  container size treatments.

\
Next, we performed a series of model simulations to test possible C allocation scenarios to account for the over predictions of seedling C mass. Testing the sensitivity of the model to observed treatment-specific LMF from the final harvest (M~s1~), which were each lower than the optmized LMF value (see Figure `r allocfig`a), improved model predictions of seedling C mass but still overestimated seedling total C by 32&pm;11.1 % (Figure `r massmodelfig2`a). Using harvest values of LMF, however, does not capture the observed increase in leaf turnover of seedlings in small containers (Figure `r allomfig`c). Thus, the use of harvest LMF values for seedlings in containers in M~s1~ likely underestimates daily C allocation to leaves over the final montsh of the experiment. Increases of 50 % in non-leaf tissue respiration (M~s2~) improved biomass estimates slightly but overestimated mass C by an average of 46&pm;9.3 % in seedlings with soil volume restriction (Figure `r massmodelfig2`b). With M~3~, non-tissue respiration rates would need to be increased by ca. 250% in order for mass balance to be achieved.

@@

\
Decreases in A~sat~ occurred at the same time as reductions in height and diameter of seedlings in containers. This initially suggests a strong link between growth and an apparent down regulation of A. However, there are several possible mechanisms that can explain reduced A in small containers including nutrient content, water or reduced sink strength [@poorter2012pot].  It was therefore necessary to examine each of these factors to determine if the induced belowground sink limitation actually triggered photosynthetic down regulation.

@@

## Conclusions
With a novel field-based design we detected a massive effect of container volume on seedling growth, not between containers but with naturally planted seedlings. This is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants has seldom been studied. Although biomass partitioning was conserved, our empirical and model results suggest that the amount of photosynthate allocated to non-biomass pools such as TNC or respiration were likely altered by sink inhibition. The debate over how rates of photosynthesis affect plant growth or to what degree these rates are instead controlled by growth has existed for decades [@sweet1966role]. Our combined empirical and modeling approach shows that when non-photosynthesis parameters were kept contant changes in A were not able to fully to predict changes in growth, an important distinction often missed in studies that manipulate source/sink activity. 
K&#246;rner [-@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control A through the demand for C. Although this may be true, we results infer that quantifying the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.
